\clearpage
\section{Indeterminates}
\defsec{indeterminates}


The indeterminates part of \LibCharSet can be separated into

\begin{itemize}
\item differential indeterminates and
\item supportive domains and categories.
\end{itemize}





\subsection{Differential indeterminates}

The differential indeterminates part consists of 
\begin{itemize}
\item \adtype{DifferentialVariableType},
\item \adtype{OrdinaryDifferentialVariableType},
\item \adtype{EliminationOrderedDifferentialVariableType},
\item \adtype{DifferentialVariable},
\item \adtype{OrdinaryDifferentialVariable},
\item \adtype{DifferentialVariableTools},
\end{itemize}

\adtype{DifferentialVariableType} is the main category for differential indeterminates.

\adtype{OrdinaryDifferentialVariableType} refines \adtype{DifferentialVariableType} for differential settings with only a single derivation.

\adtype{EliminationOrderedDifferentialVariableType} marks indeterminates to obey an elimination order.

\adtype{DifferentialVariable} is a general purpuse implementation of differential indeterminates. The file \file{dvars/dvars.as} illustrates its usage.

\clearpage
\inputSourceFile{dvars/dvars.as}{Using differential indeterminates}

\inputOutput{dvars/dvars}

\adtype{OrdinaryDifferentialVariable} implements \adtype{OrdinaryDifferentialVariableType} 

\adtype{DifferentialVariableTools} provides functions to convert betwee differential indeterminates and expression trees.


\subsection{Supportive Domains and categories}

\begin{itemize}
\item \adtype{MergedVariableType},
\item \adtype{VariableMerger}, and
\item \adtype{VariableMergerTools}.
\end{itemize}

\adtype{MergedVariableType} is a category for merging two indeterminate domains.

\adtype{VariableMerger} merges two indeterminate domains.

\adtype{VariableMergerTools} adjoins indeterminates to a given indeterminate domain. An example for \adtype{VariableMergerTools} is given in the file \file{merger/merger.as}

\inputSourceFile{merger/merger.as}{Adjoining indeterminates}

\clearpage

\inputOutput{merger/merger}
